---
title: "How I Spend My Time"
subtitle: "SAT 231: Calendar Query"
author: "Kauila Tengan"
date: "Last updated \\today"
output: pdf_document
---

```{r setup, include = FALSE}
# load packages
library(tidyverse)
library(lubridate)
library(ical)
library(kableExtra)

# set code chunk defaults
knitr::opts_chunk$set(tidy = F, # display code as typed
                      size = "small", # slightly smaller code font
                      message = FALSE,
                      warning = FALSE,
                      comment = "\t") 

# set black & white default plot theme 
# theme_set(theme_classic()) # uncomment this line if desired

# improve digit and NA display 
options(scipen = 1, knitr.kable.NA = '')
```

<!--
Please remove my comments from the code chunks and customize the code chunks on your own, including chunk labels and options and code documentation.

You can leave my markdown comments for guidance in writing your report.

Please rename the headings. 
You may add subheadings as desired to break up the report. 
-->

# Data Wrangling <!-- (e.g., How do I spend my time?) -->

<!--
The code below is provided as a starting point for importing an `ics` file into R as a data frame, doing some initial wrangling, and adding up the time spent on each activity. You are not required to use this code, so delete or modify it as you see fit. 

If you are using the code, be sure to update the capitalized parts with the relevant path to where your `ics` file is located, and set `eval = TRUE`.

As a reminder here's the link for exporting a Google calendar to ics (if on a Mac, you can hold cmd + click to open the page):
https://support.google.com/calendar/answer/37111?hl=en&ref_topic=10509946
-->
```{r Data Wrangling, warning = FALSE}
# Data import and preliminary wrangling
calendar_data <- "ktdata3.ics" %>% 
  # Use ical package to import into R
  ical_parse_df() %>% 
  # Convert to "tibble" data frame format
  as_tibble() %>% 
  mutate(
    # Use lubricate package to wrangle dates and times
    start_datetime = with_tz(start, tzone = "America/New_York"),
    end_datetime = with_tz(end, tzone = "America/New_York"),
    duration_min = difftime(end_datetime, start_datetime, units = "mins"),
    duration_hours = duration_min/60,
    # duration_min = end_datetime - start_datetime,
    date = floor_date(start_datetime, unit = "day"),
    # Convert calendar entry to all lowercase and rename
    activity = tolower(summary),
    overall = fct_collapse(factor(summary), #new overall variables for work/class
                           class = c("ASLC_class", "ASLC_study",
                                     "ENST_class", "ENST_OH", "ENST_study", 
                                     "NS_class", "NS_study", "STAT_class",
                                     "STAT_OH", "STAT_SDS", "STAT_study"),
                           work = c("work_divtern", "work_meeting", "work_NISA")
                           ),
    #making duration into a numeric
    time = as.numeric(duration_min),
    time_hrs = as.numeric(duration_hours))


  

  #data wrangling for visualization 1
  class_data <- calendar_data %>% 
     #filter our overall variable by class only
    filter(overall == "class") %>% 
    #separating class and activity
    separate(summary, c("class", "activity"), "_", remove = FALSE) %>% 
    group_by(class, activity) %>% 
    summarize(overall_time = sum(time_hrs))
   

# Compute total duration of time for each day & activity
activities <- calendar_data %>% 
  group_by(date, overall) %>%
  summarize(duration_min = sum(duration_min))


```

<!--
Describe your question(s) here.  Briefly describe your data collection process, including how you defined variables of interest.
-->

# Questions and Design



## Questions of Interest: 


**Question 1:** *How much time do I spend on each course?* 


I wanted to explore how much time I was spending on each class. Within this question, I had two aspects of interest. Looking solely at the class data in my calendar, I wanted to see how much total time was devote to each course and compare that data with the other classes. Secondly, I wanted to see the breakdown of each individual class. Within any given class, I wanted to know how much time I devoted to studying, actually attending the class, visiting office hours, and if applicable, how much time was spent attending SDS hours. 

**Question 2:** *How did my times spent on each activity during those two weeks compare to one another?*


For my second question, I wanted to look at the time devoted to all major activities during my day. I wanted to explore any interesting trends over the two weeks of data collection. I was curious how much of my day I was devoting to school, exercise, self care, leisure, sleep, and paid work. 


## Data Collection & Variables:

Going into the project, I already had an idea for my visualizations and table. I inputted all of my data into one Google calendar. I knew that I wanted to see both the breakdown within my courses, as well as the overall course data. I decided to stay consistent with my naming of each activity, paying particular attention to the capitalization or format of the letters. For each of my four classes, I abbreviated the department in all capital letters (ex. Data Science $\rightarrow$ STAT). I took it one step further for the specific activity. The capitalized abbreviation would then be followed by an underscore and then the activity in all lowercase (ex. studying for Data Science $\rightarrow$ STAT_study). When I began coding my visualizations, I realized that I only had one variable. Since my class and work inputs were highly specified, I had to first create a new variable `overall` and collapse my class and work data into `class` and `work` to address my second question. In order to create my stacked bar chart, segmented by activities, I had to then create an `activity` variable by separating `class` by the underscore. Here is a short breakdown of my variables and the data within each variable:

### Visualization 1

`ASLC` : Media History of Anime Seminar

`STAT` : Data Science

`NS` : Culture and Mental Health

`ENST` : Environmental Studies Senior Seminar

`study` : Total time spent reading, completing problem sets, and writing papers

`OH` : Office hours

`SDS` : SDS Fellows office hours

### Visualization 2 

`class` : Aggregate time spent attending all classes, office hours, and studying

`gym` : Time spent at the gym 

`rugby` : Time spent attending rugby practice

`sleep` : When I fall asleep until when I wake up 

`work` : Total time on paid work, attending club meetings, and any other meeting or work not related to academics 


<!--
Describe what information is conveyed through data visualization #1 (that you'll create below) here.
-->


# Results 


## Visualization 1: Breakdown of Total Time (Hrs) Per Class

### Discussion:

  The most noticeable thing about this visualization is just how little time was spent on my Culture and Mental Health class compared to the other three. This Hampshire College course is scheduled to meet on Tuesday and Thursday for an hour and twenty minutes each day. By absolute coincidence during these two weeks of data collection, both September 16th and September 23rd were cancelled for Yom Kippur and Hampshire Advising Day, respectively. 
  
  The distributions for my other three courses were similar and what I would have expected to observe. `ASLC` and `ENST` are both roughly three-hour seminars that meet once a week, so that class time should be greater than `STAT`. Additionally, this data shows me that I could, and should, be utilizing office hours more to reduce the overall time I spend studying. During those 16 days (384 hours), I only attended office and SDS hours a total of about two hours across those three classes. Contrary to the matter, I spent between 10 and 12 hours per class studying. Instead of struggling through problem sets, readings, and papers on my own, I would like to attend more office hours in the future.
  
### Code and Output:  

```{r Figure 1, eval = TRUE, fig.align='center'}
#creating plot from data in data frame class_data, segment the class by activity

ggplot(data = class_data, aes(fill = activity, y = overall_time, x = class)) +
  #create a stacked segmented bar chart
  geom_bar(position = "stack", stat = "identity") +
  #clean up the labels, add title
  labs(title = "Time Devoted to Each Class",
       x = "Classes",
       y = "Total Time (hrs)"
       ) +
  #add more breaks on the y axis
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10))
```

<!--
Describe what information is conveyed through data visualization #2 (that you'll create below) here.
-->

## Visualization 2: Time (Hrs) Devoted to Each Activity Per Day

### Discussion:

  Since I was tracking so many different activities, the graph is very cluttered and in particular regions, indiscernible. There are three major things that this graph describes. On average, I tend to get between seven and nine hours of sleep. However, on the night of September 20th, I had an overwhelming work load. I did work until three o'clock in the morning, but had a nine o'clock class at Hampshire College. Time spend on class (orange) increased from 3.5 hours the previous night to 5.5 hours. Since sleeping (purple) and being awake are mutually exclusive, my sleep significantly decreased from 9 hours to 4.5 hours. 
  
  The second major facet was `selfcare` and `leisure`. After the rigorous academic week, I made sure to spend the entire weekend healing. I walked around the art museum, took hikes in nature, journaled, meditated, and spent time with friends. My recorded self care on Saturday was 5.5 hours. I also noticed an interested day-by-day trend; when my class work increases, so does my leisure (green) time. I make a conscious effort to balance my workload with leisure. If I do school work for a few hours, I then reward myself with time to watch my favorite shows or to play video games that same day.
  
  The last thing that stood out to me was my inability to attend rugby (aqua) practice. Rugby is basically nonexistent on my graph. During the 16-day period, I was only able to attend two practices due to conflicts with other commitments. I was already considering quitting the team due to my lack of involvement, and this graph elucidated the fact that I simply do not have time for that particular activity. 
  
  
### Code and Output:

```{r Figure 2, fig.align='center'}

#create a time series separated by each overall activity
ggplot(calendar_data, aes(x = date, y = time_hrs,
                          color = overall)) + 
  geom_line() +
  #make the labels look like and convey accurate information
  labs(title = "Time Series Graph Over 2 Weeks",
       subtitle = "Daily Allocation of Time to Each Activity",
       y = "Total Time (Hrs)",
       x = "Date (YYYY-MM-DD)",
       color = "Overall", lty = "Overall") +
  #Change date breaks on x axis so I can see every day over the two weeks
  scale_x_datetime(date_breaks = "1 days") +
  #Increase y labels (add 10 breaks)
  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) + 
  #Rotates x axis labels 90 degrees and size for readability 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = rel(1)))
         
       
```

<!--
Describe what information is conveyed through the table (that you'll create below) here.

For nicer tables, check out:

* the kable function in the knitr package + the kableExtra package  
  https://bookdown.org/yihui/rmarkdown-cookbook/kable.html 
  https://bookdown.org/yihui/rmarkdown-cookbook/kableextra.html

or 

* the xtable package
  https://cran.r-project.org/web/packages/xtable/vignettes/xtableGallery.pdf
-->
## Table 1: Overall Summary of Each Activity 

### Discussion:

  This table describes the average, minimum, maximum, and total hours devoted to each activity. `N` also indicates the number of times that activity was recorded in my calendar. Most of my time, undeniably, was spent sleeping. 
  Due to the way I collected and wrangled my data, the average, minimum, and maximum hours fall short of conveying comparable data. These benchmarks are simply relative to the number of observations. For `sleep`, this table shows that whenever I inputted sleep data, it was for an average of 7.8 hours. On the other hand, `class` refers to any time input that was academic related. This includes times that I attended office hours for only 30 minutes, or if I had logged 40 minutes of work between classes. There were 37 recorded instances of anything school related. Even though my `Max` recorded `class` time In my opinion, the most important column in this table is `Total`. Aside from hours sleeping, I spent a significant amount of time doing anything academic related. Total `class` time was about 60 hours. I also participated in leisurely activity a lot when compared to other facets of my life. Excluding sleep, whenever I participated in a given activity, I devoted an average of about 1.5 hours to it.   
  
### Code and Output:  

```{r Table 1, results = 'asis'}
# Create new data frame from data frame calendar_data
table <- calendar_data %>% 
  # I am interested just in the overall data
  group_by(overall) %>% 
  #Renaming variable overall
  rename(Activity = overall) %>% 
  summarize(
    N = n(),
    Average = mean(time_hrs),
    Min = min(time_hrs),
    Max = max(time_hrs),
    Total = sum(time_hrs)
    ) %>% 
  #arrange in descending order
  arrange(desc(Average)) %>% 
  #make aesthetically pleasing and change number of digits following decimal.
  kable(booktabs = TRUE, digits = 1, caption = "Activity Summary Statistics (Hrs)") %>% 
  #center and change font size
  kable_styling(font_size = 15) 

#output table
table
```

<!--
To conclude, briefly summarize what you found in response to the questions posed here.
-->


\newpage
# YOUR REFLECTION HEADING <!-- e.g., Reflection -->

<!--
Write your one-page reflection here in paragraph form.  In particular, address:

* What difficulties in the data collection and analysis process did you encounter?  Identify two of your main hurdles in gathering accurate data.

* What implications does that have for future data collection and/or analysis projects? 

* How much data do you think you'd need to collect in order to answer your question(s) of interest? Would it be hard to collect that data? Why or why not?

* As someone who provides data, what expectations do you have when you give your data (e.g. to Facebook, Google, MapMyRun, etc.)?  

* As someone who analyzes others' data, what ethical responsibilities do you have?
-->


